---
title: "Exploring Race in Prisons"
output:
  html_document:
    df_print: paged
---

```{r include = FALSE}
source("analysis.R")
```

### Introduction 
The criminal justice system of America is no stranger to scrutiny. Plagued with detrimental biases within its computer systems and people, we are all too familiar with the topic of race and prison. Using the VERA institute's dataset, I have chosen to focus on **prison** statistics at the state and county level, analyzing **race** factors. Specifically, the variables I mainly focused on were: `_prison_pop`, `_prison_adm_rate`, and `prison_pop_rate`. I focused on California and New York for my plots. 

### Summary Information 
Here are some quick summary facts I've extracted from my analysis: 
- Cali saw the highest number of Black prison population in **2008**, at 49464 people
- The highest recorded number of Black prison population in a county was New York County, at 24818 people
- The year of the most 
- The highest black to white proportion in 2016 was at an estimated 1915% ! 
- The state with the highest ever Black to white proportion was in New Jersey 

### Trends over Time Chart 
Though my focus was on the Black prison population, I decided to also map out the varying races. 
The plot is divided by 2 different states: California and New York. Then, the prison population is compared for the different groups of race, over years. I wanted to see 

```{r echo = FALSE, warning=FALSE}
prison_year_pop_plot
```

